---
title: "project ivy: House Prices (Kaggle) - EDA"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(knitr)
source('src/ini.R')
source('src/load.R')
source('src/transform.R')
```

## Overview


[House Prices: Advanced Regression Techniques](https://www.kaggle.com/c/house-prices-advanced-regression-techniques)

## Let's get a feel for the data

Let's load the data and implement the rules based on **data_description.txt** and inspect the data

```{r overview, echo=FALSE, warning=FALSE}
# dimensions
# -----------
dims = dim(df)
names(dims) = c('Rows', 'Cols')
print(dims)

# NA's
# -----------
print_na_cols = function(){
  na_cols = lapply(df, function(x){sum(is.na(x))}) %>%
  unlist() %>%
  as.data.frame() %>%
  rownames_to_column()
names(na_cols) = c('Variable', 'Missing')
na_cols %>%
  filter(Missing > 0) %>%
  print()
}
print_na_cols()
```

There are 21 variables containing NA's. Note the following:

*  **SalesPrice** is missing because the **test** data was merged with the **train** data. That leaves us with 20 variables containing NA's.
*  **GarageYrBlt** will be NA if there is no garage on the property

Let's visualise missing
```{r impute_viz, echo=FALSE, warning=FALSE, message=FALSE}
require(VIM)
df %>%
  select(-GarageYrBlt, -SalePrice, -Id, -set_id) %>%
  aggr(.)

```


```{r imputed, echo=FALSE, warning=FALSE}
source('src/impute.R')
print_na_cols()
```

## Visualisation

Notice **SalePrice** vs **SalePriceLog**. We'll be using **SalePriceLog** from now on.

```{r viz, echo=FALSE, fig.width=10, fig.asp=.25}
train = df %>%
  filter(set_id == '1')
test = df %>%
  filter(set_id == '2')
train$SalePriceLog = log10(train$SalePrice)

p = ggplot(train) + theme_bw()
multiplot(p + geom_histogram(aes(x = SalePrice)),
          p + geom_histogram(aes(x = SalePriceLog)),
          cols = 2)

```

Box plots

```{r, fig.width=10, fig.asp=.25, echo=FALSE}
draw_boxplot = function(var){
  p + geom_boxplot(aes_string(x = var, y = 'SalePriceLog')) + ggtitle(var)
}

bp_vars = c(factor_vars, ord_factor_vars)
plotlist = lapply(bp_vars, function(x){draw_boxplot(x)})
```
```{r, fig.width=10, fig.asp=.75, echo=FALSE}
multiplot(cols = 4, plotlist = plotlist[1:9])
```
```{r, fig.width=10, fig.asp=.75, echo=FALSE}
multiplot(cols = 4, plotlist = plotlist[10:18])
```
```{r, fig.width=10, fig.asp=.75, echo=FALSE}
multiplot(cols = 4, plotlist = plotlist[19:27])
```
```{r, fig.width=10, fig.asp=.75, echo=FALSE}
multiplot(cols = 4, plotlist = plotlist[28:36])
```
```{r, fig.width=10, fig.asp=.75, echo=FALSE}
multiplot(cols = 4, plotlist = plotlist[37:44])
```


